Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
clean code and update text in 2.2

(This section discusses the tests that will be used in the analysis and shows the results for the residual plots displayed in the previous section.)

Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depending on the alternative hypothesis that is focused on, a variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. And for testing normality, the Shapiro--Wilk test [ref here] is perhaps the most widely used test included by many of the statistical softwares. Another choice will be the Jarque--Bera test [ref here] which directly checks if the sample skewness and kurtosis match a normal distribution.

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

```{r}

mod <- poly_model(include_z = FALSE, sigma = 0.25)
dat <- mod$gen(300)

dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat, x, e))
dat2 <- heter_model(b = 64)$gen(300, computed = select(dat, x, e))
dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
               gen(300, computed = select(dat, x))

data.frame(plot = c("A", "B", "C", "D"), 
           model = c("None", "Non-linearity", "Heteroskedasticity", "Non-normality"),
           f = c(POLY_MODEL$test(dat)$p_value, 
                 POLY_MODEL$test(dat1)$p_value,
                 POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
                 POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
           b = c(HETER_MODEL$test(dat)$p_value, 
                 HETER_MODEL$test(dat1)$p_value,
                 HETER_MODEL$test(dat2)$p_value,
                 HETER_MODEL$test(dat3)$p_value),
           s = c(shapiro.test(dat$.resid)$p.value, 
                 shapiro.test(dat1$.resid)$p.value,
                 shapiro.test(dat2$.resid)$p.value,
                 shapiro.test(dat3$.resid)$p.value)) %>%
  mutate(f = format(round(f, digits = 3), nsmall = 3), b = format(round(b, digits = 3), nsmall = 3), s = format(round(s, digits = 3), nsmall = 3)) -> table_dat

  table_dat$f <- kableExtra::cell_spec(table_dat$f, italic = as.numeric(table_dat$f) < 0.05)
  table_dat$b <- kableExtra::cell_spec(table_dat$b, italic = as.numeric(table_dat$b) < 0.05)
  table_dat$s <- kableExtra::cell_spec(table_dat$s, italic = as.numeric(table_dat$s) < 0.05)

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "RESET", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" )

```

@@

(This section introduces the lineup protocol, and briefly discusses the method for sampling null data, calculating the p-value and estimating power.)

@@

Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive. Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

>>>>

Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depending on the alternative hypothesis that is focused on, a variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. And for testing normality, the Shapiro--Wilk test [@shapiro1965analysis] is perhaps the most widely used test included by many of the statistical softwares. Another choice will be the Jarque--Bera test [@jarque1980efficient] which directly checks if the sample skewness and kurtosis match a normal distribution.

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject the null hypothesis due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

```{r}
set.seed(10086)

# Base data
dat0 <- mod$gen(300)

# Replicate data in Figure 1
dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
               gen(300, computed = select(dat0, x))

dat_list <- list(dat0, dat1, dat2, dat3)

table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
                        model = c("None", 
                                  "Non-linearity", 
                                  "Heteroskedasticity", 
                                  "Non-normality"),
                        r = map_dbl(dat_list, 
                                    ~POLY_MODEL$test(.x, 
                                                     test = "RESET", 
                                                     power = 2:4)$p_value),
                        b = map_dbl(dat_list, 
                                    ~HETER_MODEL$test(.x)$p_value),

                        # SW test has not been included by visage
                        s = map_dbl(dat_list, 
                                    ~shapiro.test(.x$.resid)$p.value)) %>%

  # 3 decimal points
  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%

  # Italic if reject
  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) < 0.05)))

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "Breusch-Pagan", 
                             "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

@@

@@

One may prefer conventional hypothesis testing because reading diagnostic plots is to some extent subjective and indecisive as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984]. It is arguably true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
clean code and update text in 2.2

(This section discusses the tests that will be used in the analysis and
shows the results for the residual plots displayed in the previous
section.)

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing for detecting model defects. Depending on the
alternative hypothesis that is focused on, a variety of tests can be
applied. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. For
testing non-linearity, one may apply the F-test as a model structural
test to examine the significance of specific polynomial and non-linear
forms of the regressors, or the significance of proxy variables as in
the Ramsey Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
test {[}ref here{]} is perhaps the most widely used test included by
many of the statistical softwares. Another choice will be the
Jarque--Bera test {[}ref here{]} which directly checks if the sample
skewness and kurtosis match a normal distribution.

Example residual plots given in Figure
\ref{fig:residual-plot-common-departures} are examined by the
corresponding model structural test, Breusch-Pagan test and
Shapiro--Wilk test as shown in Table
\ref{tab:example-residual-plot-table}. In the example, both the
Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis for departures that they do not intend to examine. As
discussed in \citet{cook1982residuals}, most residual-based tests for a
particular type of departure from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of the null hypothesis despite when majority of the residuals
are well-behaved \citep{cook_applied_1999}. This can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers.

@@

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.434 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.378 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.667 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}

>>>>

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing for detecting model defects. Depending on the
alternative hypothesis that is focused on, a variety of tests can be
applied. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. For
testing non-linearity, one may apply the F-test as a model structural
test to examine the significance of specific polynomial and non-linear
forms of the regressors, or the significance of proxy variables as in
the Ramsey Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
test \citep{shapiro1965analysis} is perhaps the most widely used test
included by many of the statistical softwares. Another choice will be
the Jarque--Bera test \citep{jarque1980efficient} which directly checks
if the sample skewness and kurtosis match a normal distribution.

Example residual plots given in Figure
\ref{fig:residual-plot-common-departures} are examined by the
corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as
shown in Table \ref{tab:example-residual-plot-table}. In the example,
both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis for departures that they do not intend to examine. As
discussed in \citet{cook1982residuals}, most residual-based tests for a
particular type of departure from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of the null hypothesis despite when majority of the residuals
are well-behaved \citep{cook_applied_1999}. Furthermore, with a
sufficiently large sample size, residual-based tests may reject the null
hypothesis due to a slight departure that is of little practical
significance. These can be largely avoided in diagnostic plots as
experienced analysts can evaluate the acceptability of assumptions
flexibly, even in the presence of outliers and slight departures.

@@

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.779 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.658 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.863 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}
